---
title: "Getting and Cleaning Data Final Project - Draft 1"
author: "Andrew Mauro"
date: "January 24, 2017"
output: html_document
---

1. Load Data and Merge the Training and the Test sets to create one data set
============================================================================

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


rm(list=ls())
#packages
library(data.table)
library(dplyr)

#train, test, names, and subjects
train <- fread("UCI HAR Dataset/train/X_train.txt", sep = "auto")
train_label <- fread("UCI HAR Dataset/train/y_train.txt", sep = "auto")
test <- fread("UCI HAR Dataset/test/X_test.txt", sep = "auto")
test_label <- fread("UCI HAR Dataset/test/y_test.txt", sep = "auto")


names <- fread("UCI HAR Dataset/features.txt",sep="auto")
subject_train <- fread("UCI HAR Dataset/train/subject_train.txt", sep = "auto")
subject_test <- fread("UCI HAR Dataset/test/subject_test.txt", sep = "auto")
labels <- fread("UCI HAR Dataset/activity_labels.txt",sep="auto")

#apply names to objects
names(test) <- names$V2
names(train) <- names$V2
names(train_label) <- "activityid"
names(test_label) <- "activityid"
names(subject_train) <- "subjectid"
names(subject_test) <- "subjectid"

#append subject labels and activity labels to train and test
train <- cbind(subject_train, train, train_label)
test <- cbind(subject_test, test, test_label)

#bind train and test and properly format names
bound <- data.frame(rbind(train,test))
names(bound) <- make.names(names(bound),unique = TRUE)

#remove
rm(names, subject_test, subject_train, train, test, labels, test_label, train_label)
```


**Conclusion:** We now have one data set, 'bound', a 10,299 by 562 data frame with our training and test data. We have not appended activity labels yet.

2. Extract from 'bound' the measurements on the mean and standard deviation for each measurement
===============================================================================================

```{r}
#select only variable names with -mean. and -std.
bound <- bound[,grep("[Mm]ean|std|activityid|subjectid", names(bound))]
```

3. Replace Activity labels with descriptive labels

```{r}
bound$activityid<- factor(bound$activityid, labels = c("walking", "walking upstairs", "walking downstairs", "sitting", "standing", "laying"))
```


4. Label data set variables with descriptive variable names

```{r}
names(bound) <- gsub("\\.", "", names(bound))
names(bound) <- gsub("^t", "TimeDomain", names(bound))
names(bound) <- gsub("tBody", "TimeDomainBody", names(bound))
names(bound) <- gsub("Body", "Body", names(bound))
names(bound) <- gsub("Acc", "Acceleration", names(bound))
names(bound) <- gsub("Gyro", "Gyroscope", names(bound))
names(bound) <- gsub("Mag", "Magnitude", names(bound))
names(bound) <- gsub("^f", "FrenquencyDomain", names(bound))
names(bound) <- gsub("angle", "Angle", names(bound))
```

5. Create an independent tidy data set with the average of each variable for each activity and each subject

```{r}
tidy <- tbl_df(bound)%>%
                group_by(activityid, subjectid)%>%
                summarise_each(funs(mean))
tidy
write.table(tidy, "tidy.txt", row.names = FALSE, sep = "\t")


```
